---
title: "Master's Thesis"
author: "Michael Gaffney"
date: "January 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, error=F, warning=F, fig.width = 10, fig.height = 6)
source('analysis2.R')
library(patchwork)
```

## Introduction
Long considered the result of harmful byproducts or malfunctions, depression, crying, self-harm, suicide attempts, and illness symptoms have all gained increased attention for their potential to function in part as signals of need or ways of bringing about positive changes in social relationships. Although based mainly in signaling theory, such approaches have benefited greatly from economics and game theory, with the result being explanations for the existence of such states that emphasize the causal role of one’s social environment in addition to one’s internal state. From this perspective, all the proposed signals are expected to share many meaningful similarities due to their shared function, with their differences stemming largely from differences in the types of need being experienced or the target audience.

Likely the main reason why these conditions are predominantly thought of as maladaptive is the harm they cause the one experiencing them, which can be quite severe with the costlier signals (Kessler 2012). However, subjective suffering and negative fitness outcomes are not inextricably linked. This is most obviously seen with negative emotions motivating fitness enhancing behavioral shifts, as is widely recognized to be the case with hunger, thirst, or pain (Nesse 1990; Tooby and Cosmides 2008). Like with negative emotions, the symptoms associated with the proposed signals of need may not simply be the result of mistakes. Instead, they are proposed to have a functional role in increasing the confidence of others who may be motivated to help that one’s need is genuine through the costs involved.

To this point, the evidence suggesting that the proposed signals of need are adaptive comes from outside indications of increased social support. For example, major depression, suicide attempts, placebo responses, and crying are all more common than should be expected of maladaptive malfunctions due to the likelihood of strong selection pressures against those who incur extra costs in times of need (cite). For the proposed signals of need to still be maladaptive with this prevalence, the traits responsible for such malfunctions would likely have to result primarily from environmental mismatches (Hidaka 2012) or pleiotropic traits in which major benefits come from the genes responsible for the proposed signals of need (Williams 1957). However, both neither situation is safe to assume due to the proposed signals of need being found across a wide range of environments and social settings (Bromet et al., 2011; Syme et al. 2016), and the fact that genetic modularity often results in the potential of other genes to limit the negative effects of pleiotropy while preserving the benefits in polygenic traits (Hansen 2003). 

Another important indication that the proposed signals of need are not likely to solely be malfunctions is the fact they tend occur when individuals are in need and can benefit from others. This is seen in both depression and suicide attempts occurring most frequently in times of social adversity in which individuals are experiencing costs due to the actions of others or the loss of existing social relationships (Hammen 2005; Mazure 1998; Syme et al. 2016). Similarly, placebo responses, which have been proposed to be the downregulation of signaling through increased illness symptom severity (Steinkopf, 2015; Tiokhin, 2016), often happen in response to individuals receiving support after feeling a lack of it in conjunction with perceived illness (Humphrey 2002). In both situations, individuals can greatly benefit from the help of others due to the fitness costs that accompany social adversity and illness (Syme et al. 2016; Sugiyama and Sugiyama 2003).

In this thesis, I report the findings of a vignette study conducted with Edward Hagen examining if the proposed signals of need result in positive responses as predicted by certain signaling explanations. This involved exposing participants to a fictional scenario in which an imagined sister requested substantial support from them under various social contexts. In addition to the signals themselves, the amount of conflict within a relationship and the degree to which one had information about whether or not the person in the vignette is genuinely in need were manipulated, as both types of information are expected to be important inputs for our adaptations responsible for helping others in need.

## Social support in humans

Such a scenario in which costly emotional states lead to increased benefits is possible due to the degree to which humans cooperate with another, something which was necessary for our survival for much of our evolutionary history and is reflected in our adaptations today (Sugiyama and Sugiyama 2003). Unlike most species, this support can often come from a wide range of individuals in humans, with kin beyond parents playing an important role in childcare across generations (Kaplan 2009; Meehan 2005; Sear and Mace 2008) and unrelated individuals often helping in a wide range of situations due to the benefits of both direct and indirect reciprocity as well as reputation gains (Jaeggi and Gurven 2013; Barclay 2013).

While important at almost any time, cooperation is especially important in times of need. Illness, injury, and physical conflict with others have all been recurrent problems over the course of human evolution, with their outcomes often resulting in times in which individuals could not survive without short-term support (Sugiyama and Chacon 2000). In these times, help often takes the form of temporary increases in provisioning or labor to decrease the workload of the ill (Sugiyama and Sugiyama 2003), with the existence of long-term support of those who are unlikely to recover in both modern humans and Neandertals (Trinkaus and Villotte 2017) clearly showing that such cooperation does not rely solely on expectations of direct reciprocity (Sugiyama and Sugiyama 2003) Although rarely examined empirically, both the fact that humans can reliably find support in times of need and the presence of long-term support suggest that lasting changes might be possible in times of need outside of instances of disability.

Despite the likelihood of individuals benefiting from increasing their support of others in times of need and widespread evidence that it often happens, devoting a large amount of resources to another is only likely to be worth it if the need is genuine. It is in these instances when the proposed signals of need are expected to be adaptive, so long as the result is an increase in the likelihood of support due to the signal enhancing one’s appearance of need and the benefits outweighing the costs of producing the signal (Zahavi 1993). This limitation precludes the costlier signals from being adaptive in relatively minor instances of need but does rule out adaptive explanations in times when one’s need is great and unlikely to be believed.

## Costly Signaling

All organisms can benefit from altering the behavior, physiology, and psychological states of others, something which is seen from single celled organisms to social animals (Dawkins 1982). Such changes can be accomplished through chemical or behavioral means, with information flowing from the signaler to the receiver in ways meant to benefit the one sending the message, regardless of whether or not it benefits the one receiving it (Dawkins 1976).

In social animals, members of one’s own species are often the target, with sexual selection and intrasexual competition creating many situations in which one’s reproductive success depends greatly on the responses of others. The results of this are strong selection pressures favoring those who can make the most of these opportunities to outcompete others and receive the largest benefits through cooperation or aggression. 
Of the strategies used to alter the behavior of conspecifics, costly signaling to provide believable information about one’s quality or current state is likely to be one of the most common and effective ways to do so (Zahavi 1975). A signal differs from a cue in that it is intentionally sent by an organism to alter the behavior or thoughts of other individuals, while a cue is any piece of information an organism uses to alter its behavior or physiological state that was created without the intent to do so (Maynard Smith and Harper 1995). Since signals are intentionally created, they are easy to fake due to the fact they can be disconnected from the qualities being signaled (Zahavi 1975). This coupled with the selection pressures resulting from the fact that conflicts of interest are ubiquitous among living things (Trivers 1974; Zahavi 1993) means that there is much to gain from successfully faking a signal as one could receive the benefits without paying the costs of an honest display. Therefore, strong selection pressures exist favoring adaptations to spot fake signals, something which has been shown to exist in many species (Hasson 1994). 

To overcome such counter-adaptations to cheating, an organism may produce a signal that comes with costs to create a signal that is more likely to be believed by the targets of the signal (Zahavi 1975). Such costly signals work because they are only worth producing if one has the underlying qualities associated with the signal (Zahavi 1993). The result of this is a relatively honest display compared to less costly signals and those more disconnected from the quality being displayed. This is most commonly examined in instances of signaling attractiveness, where the costs associated with physical weaponry or conspicuous traits result in lowered chances of survival in all that display them but relatively prohibitive costs being experienced by those who cannot afford the honest display (Zahavi 1975; Gangestad and Scheyd 2005). When it comes to the proposed signals of need, costlier signals can be thought of as honest as only those who are experiencing need can benefit from displaying them, as their costs are prohibitively high for those who are not in need. 

## Depression

Of all the proposed signals of need, major depression has received the most attention and likely causes the most suffering. Although often overlooked due to its indirect connection with mortality (Patel et al., 2004), it is the leading cause of nonfatal disease burden in the world and the second greatest contributor to global years lost to disability (Ferrari et al., 2013; Slavich & Irwin, 2014). Around 8% of people in the U.S. experience major depression each year, and almost 20% will experience major depression in their lifetime (Kessler et al. 2010). Although exceptions exist, similar prevalences are often reported for depression as a whole, with there being some evidence that it is more often diagnosed in developed countries (Bromet et al., 2011). However, in no country is there a good understanding about how people respond to the depressed.

Major depression typically involves a large-scale reduction in activity along with persistent negative feelings (Fried and Nesse 2014). These symptoms take many forms but generally fall into the categories of insomnia, weight loss, fatigue, agitation, attention problems and feelings of worthlessness. As these symptoms can accompany normal sadness, length is primarily used to distinguish it from minor depression, with DSM-5 requiring the existence of at least five symptoms for at least two weeks (American Psychiatric Association, 2013). However, it often also differs in severity, as is seen with depression’s connection to increased suicidal ideation (Bostwick and Pankrantz 2000). From an evolutionary perspective, it can be described as partial to complete abandonment of normally fitness enhancing behavior that also results in a reduction of benefits one provides to others (Hagen 2003).

Most cases of depression are closely tied to major negative life events (Hammen, 2005). This is clearly seen when comparing those with depression other groups, with depressed individuals often reporting at least twice as many instances of negative events than nondepressed individuals (Mazure 1998) and more negative events than those mental disorders like schizophrenia and bipolar disorder across a wide range of studies (Paykel 1994).  In addition to differences in recent experience, the connection between negative events and depression is also seen in the high percentage of depression cases that are preceded by negative event and the rapid onset of depression following the event, both of which are seen in a largescale study of female twins that found 80% of depression cases were preceded by at least one negative life event and symptoms tending to occur within one month of the event (Kendler et al. 1999). 

This strong connection between negative life events and depression along with the need and distress that likely accompanies these events has led to the suggestion that this relationship is causal (Kendler et al. 1999). Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one’s control, which shows that the connection is unlikely to be driven solely by depressed individuals creating situations where negative events are likely to be common (Hammen 2005, and Kendler et al. 1999.). In addition to this, twin studies have shown that one’s history of negative events is still a strong predictor when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, with divorce being an example of this (Kendler et al. 1999; Kenlder and Baker 2007).

Of the many types of negative life events associated with depression, those which involve social adversity tend to be more strongly associated with depression (Kendler et al. 1998). Like the relationship between negative events as a whole and depression, this relationship is also seen cross-culturally, with the loss of a loved one, threat of divorce, or abusive relationships all increasing one’s chances of becoming depressed (Kendler et al., 1999, 1995). Of these predictors, physical assault has been found to be the strongest predictor of depression in the few studies in which it has been examined (Kendler et al. 1995; Kendler et al. 1999) and may have been much more common throughout our evolutionary history than is often thought based on rates of spousal abuse in small-scale societies (Stieglitz 2011). 

Conflict is likely to often be present in many of the previously mentioned forms of social adversity outside of the loss of a loved one and is often reported to be linked with depression when it is examined as a predictor (Asgeirsdottir 2011. Breakups in both romantic and cooperative relationships are often preceded by conflict (Amato and Marriot 2007), and surely contain it when one side wants reconciliation and the other does not. Likewise, abusive relationships are bound to be conflictual, with one side holding the power to harm the other and doing so. When looking solely at conflict as a predictor, both conflict without abuse and conflict with family members are also strong predictors of depression in Western societies (Kendler et al., 1995), with conflict outside of the family also being a predictor of depression among the Tsimané (Stieglitz et al., 2015). 

## Nonadaptive explanations

Explanations for the relationship between social adversity and depression often involve the malfunction of mechanisms relating to low mood, which is almost universally considered adaptive when it functions properly (Hagen, 2003; Nesse, 1999; Wakefield & First, 2012). One influential version of this is the stress-diathesis model, which suggests that stressful situations interact with one’s predisposition to depression to bring about depressive symptoms (Monroe & Simons, 1991). In general, such approaches do not make distinctions between different types of stress and their effects on risk of depression nor do they predict different outcomes based on variables outside of stress levels and personality differences. From this perspective, the social adversity often associated with depression may simply be one way for extreme levels of stress to accumulate thereby turning what would normally be short-term sadness into long-term depression. 

Such approaches also generally do not make predictions on the types of responses the depressed are likely to receive, although there are exceptions. This is due in large part to the previously mentioned costs involved, which shift the focus of most research towards treatment (Patel et al. 2004) and increases the size of the benefits needed for it to adaptive. When predictions are made about responses, they tend to involve depression resulting in negative responses due to strain the depressed’s anhedonia puts on relationships (Coyne 1976). From these perspectives, the costs of depression are simply too great to warrant a focus on the positive responses or too unexpected to be examined.

## The bargaining model of depression

Unlike these approaches, the bargaining model of depression emphasizes both the roles of social adversity and potential benefits in its explanation for why most instances of depression happen (Hagen, 2003). It views depression as an oftentimes adaptive response to social adversity or other hardships in which support from others or more beneficial cooperative relationships can be gained. From this perspective, maladaptive cases of depression are still expected to exist as any biological mechanism can fail (Hagen 2011). However, in many instances, it is expected to serve as a viable option to elicit more social support for those who may have few other options.

The bargaining model of depression is based on the idea that for much of our evolutionary history the small groups humans lived in created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop (Hagen 2003. In these instances, one’s monopoly on one or more types of cooperation could be leveraged to change how one is treated if others value the depressed’s cooperation enough and view the depressed as being in honest need if such cooperation is withdrawn. In other words, depression could function similarly to a labor strike, demonstrating that one’s need was genuine; and that if the target of the signal wanted to continue receiving benefits from the depressed, he or she would need to increase support towards the signaler. 

Like with the other proposed signals of need, the costs experienced by the depressed play an important role in its function as an indicator of honest need. As previously mentioned, depression often involves one stopping investment in oneself to severe levels. This means that depression is unprofitable to fake as the costs involved are likely only worth experiencing by those who are experiencing high levels of need, thereby increasing the confidence of others that the signal is honest as is suggested with other costly signaling hypotheses. The result is the message that due to one’s state of need “there is little difference in the fitness benefits obtained from investing heavily in her current social strategy or investing little” (Hagen 2003).

Despite predicting many cases of depression to stem from situations in which it was adaptive ancestrally, it does not suggest that depression should go untreated. This is because adaptation is not to be confused with subjective wellbeing, which is a much more important concern in every area except for explanations of why depression exists (Cosmides and Tooby 1999). However, it does suggest an explanation for why most medical treatments perform similarly to cognitive behavioral therapy (Hollon et al. 2005) and placebos (Walsh et al. 2002) in that all three methods demonstrate the existence of a knowledgeable medical professional giving active support to the depressed. In other words, such treatments may provide sufficient cues that the signal has worked, even though none may treat the underlying cause.  

## Other proposed costly signals of need

As previously mentioned, the bargaining model of depression is not unique as an oftentimes adaptive explanation for harmful symptoms, with crying (Hendriks et al., 2008; Lummaa, 1998), self-harm (Hagen et al., 2008), suicide attempts (Rosenthal, 1993; Syme et al., 2016), and illness symptoms (Steinkopf, 2015; Tiokhin, 2016) having all been argued to signal need. 

Of these proposed signals, self-harm and suicide have the most in common with depression. Both are much more common for their costs than should be expected for malfunctions (Hagen et al. 2008), with suicide attempts being found across a wide range of cultures (Syme et al. 2016; Weissman et al. 1999). As with depression, self-harm and suicide attempts tend to come in times of social adversity like divorce or neglect, both of which are situations when it is likely that social support is both greatly needed and difficult to obtain (Hagen et al. 2008). This connection to social adversity is clearly seen among instances of suicide attempts in the HRAF database, in which 77.4% of attempts were found to be linked to times of social conflict, and 88.7% were linked to fitness threats in general (Syme et al. 2016). Importantly, 84.9% of attempts were linked to situations in which attempters were powerless to resolve their situation alone (Syme et al. 2016), further indicating its potential as a signal of need in times when other less costly methods of bargaining are less likely to be successful, assuming the benefits are large enough on average.  

Unlike depression, where the costs are directly felt by the signaler and his or her targets, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. Therefore, the way others view and respond to suicide attempts are likely to be strongly influenced by the probability that the attempt would have been successful (Rosenthal 1993). Differences in these probabilities result primarily from the method of the attempt, which can range from little to no risk to almost certain death (Weiss 1957).  When viewed this way, the best candidates for bargaining are those in which there is a “non-negligible” but nonguaranteed chance of death (Rosenthal 1993). 

From this perspective, the threat of suicide alone is unlikely to result in increased support unless there is other evidence that the threat is credible, something which few things outside of an attempt can provide. However, attempts need to only be risky enough to trigger important, beneficial changes and should be predicted to closely track this threshold if it is the result of adaptations. Therefore, a key prediction of the bargaining model of suicide is that they should often result in the attempter surviving (Syme et al. 2016). This is clearly the cases across the world, with attempts that do not result in death far outnumbering those that do (CDC 2004 in Syme et al.) and individuals often choosing less risky forms of suicide when methods with higher probabilities exist (i.e. overdosing on pills instead of jumping off of a tall building) (Rosenthal 1993). 

Of the proposed signals, crying has the longest history of being examined as a signal which often results in benefits and is the least controversial due to its low costs (Hendriks et al. 2008). Infant crying is almost always considered adaptive, although how it functions to elicit support is still debated, with many proposed explanations involving mechanistic pathways outside of signaling honest need like crying serving to demonstrate quality, blackmail parents through increased risk of predation, or manipulate parents into investing less in other offspring (Lummaa et al. 1998). However, like the other proposed signals of need, it is more common when one is in need and can benefit from the support of others. This is seen both observationally, with crying bouts being shorter in cultures which have higher levels of parent-offspring contact (Lummaa et al.1998) and experimentally, with longer crying bouts being associated with shorter holding times (Hunziker and Barr 1986). 
Like infant crying, adult crying is a human universal, with emotional tears likely being unique to humans (Hendriks et al. 2007). Due to differences in what motivates support as well as differences in both the prevalence and the costs involved, it is likely that adult crying often differs from infant crying in both function and the psychological mechanisms. However, in both infants and adults, crying is an effective signal that something is wrong, with both tear production and the facial sadness that often accompanies crying both increasing the perception of sadness by others (Hasson 2009), something which and humans being well equipped to distinguish tears of joy from tears of sadness (Hendriks et al. 2007). 

One key difference between adult and infant crying is that adult crying often involves a referent, with much of adult crying stemming from a handful of particular types of situations (Borquist 1906). This means that it is likely a better candidate for a costly signal of need as envisioned here since some preexisting cost is expected to exist before the signal. In addition to once again suggesting that crying is an oftentimes adaptive response to times of need, the fact that tears of sadness rarely occur outside of times of need like loss or conflict hints that it is often an honest signal. In line with this reasoning, the absence of an effective sadness display has been shown to reduce supportive behavior, suggesting that properly displaying sadness (with which crying helps) increases the confidence of others that one is in need (Diminich dissertation citing Barr-Zisowitz 2000). However, it is unclear if costs of crying are significant enough to result in honest displays (Hasson 2009).

The costs of crying are felt across a wide range of domains, with it being hard to tell the relative impact of one cost to the next. However, crying is almost certainly a less costly signal than depression, self-harm, or suicide attempts. One important area in which the costs of crying are likely to be felt is through reputation hits, with crying often lowering how others view a person (Hendriks et al. 2008). While true of both sexes, this may be particularly true if the target of the signal is a man, as men have been shown to simply avoid crying people more often than women in experimental settings. Although not directly linked to crying, sadness, which almost certainly accompanies most instances of crying, tends to slow both cognitive and motor functioning (Ekman, 2003; Ellgring & Scherer, 1996 in Diminich dissertation) thereby imposing at least some opportunity costs surrounding and during bouts of crying, something that may be directly seen with the blurred vision that comes with tears (Hasson 2009). When sadness is already present, crying may simply exacerbate this distress, as is seen physiologically with the increased hearts rate and blood pressure that typically accompanies crying taking time to return to normal levels after crying, with their being little to no emotional benefits to crying in most instances (Hendriks et al. 2007).

## Commonalities shared by the signals of need

Although each theory emphasizing the ability of negative subjective symptoms to elicit support differs somewhat to fit the symptoms of the hypothesized signal, it is clear that they share many similarities, both empirically and theoretically. At the most basic level, all are predicted to result in benefits that make up for the costs involved, at least on average. This means that smaller forms of support are likely to be sufficient for the less-costly signals, with the costlier signals requiring larger gains all else being equal. Therefore, it is reasonable to expect the less costly signals to be more common, which is seen with instances of crying far outnumbering cases of depression, which far outnumbers the cases of suicide attempts (van Hemert et al. 2011; Kessler & Bromet, 2013; Nock and Kessler 2006).

One way to reduce the average costs of the proposed signals is to only engage in them in situations where large payoffs are relatively likely (Rosenthal, 1993). Although little has been done to investigate the existence of such mechanisms, it is a reasonable trait to expect of the proposed costly signals of need if they are adaptive. This is because the absence of such systems would decrease the amount of ways the proposed signals could have evolved as adaptations due to the role selection would play in disfavoring strategies that led to dangerous misfires and worsened conditions in times in which such outcomes may be the most costly. Without these mechanisms, the adaptive nature of the proposed signals of need may be confined to times of extreme need, as such situations may reduce the costs of signaling to such an extent that the relevance of the likelihood of support to levels where it becomes much less relevant (Hagen et al., 2008; Tiokhin, 2016).

In addition to not engaging in a signaling strategy that is likely to result in insufficient support, only investing in a signal of need when one does not have any less costly alternative options is another likely way to reduce the average costs of the proposed signals. This prediction is clearly built into the unified model of anger and depression, which proposes that anger and depression may be complementary strategies, with aggression being favored by those who have the formidability to bargain effectively through the use of force and depression favored by those that cannot (Hagen & Rosenström, 2016). Support for this comes from the finding that differences in upper body strength account for much of the female biased sex difference seen in depression. Importantly, this sexually dimorphic pattern of women experiencing depression symptoms more often than men (Hagen and Rosenström 2016) is also seen in all the other proposed costly signals examined here (van Hemert et al. 2001, CDC 2004), suggesting that less costly bargaining methods are reliably favored, assuming there is a similar scarcity of alternatives for women relative to men.

Although not derived from the similarities of the theories explored here in the same way as the previous traits, a key similarity shared by the signals is that the costs associated with signaling will almost always be less than the costs of the symptoms themselves since some level of symptomology is likely to be present that is not tied to a signaling function. In other words, the cost of engaging in a costly signal of need is often going to be significantly less than the symptom the recipient sees. This is best seen with illness symptoms, much of which are the result the body’s defense mechanisms or a pathogens circumvention of these defenses (Nesse and Williams 1995). These costs would be maintained without a signaling function; however, they can also be increased to better convince others of the signaler’s need if the situation is right (Steinkopf, 2015; Tiokhin, 2016). In this case, the costs of signaling may only be a small part of the totality of the symptom. Although such a situation is harder to imagine with depression or suicide attempts, extreme cases of conflict or abuse may result in situations in which the difference between a depressed and a nondepressed state may be much closer than normal (Hagen 2003). 

## Benefits

One of the most foundational predictions of these signaling theories is that the signaler should often receive benefits after the signal, either in the form of support or renegotiated social relationships. This is possible due to adaptations we possess for helping others which deliver benefits to the helper through kin selection, indirect reciprocity, or reputation gains (Hamilton 1964; Barclay 2013). Such support comes at a cost but is often worth it if the person in need is important to the potential helper and in a severe but likely to be overcome situation. In addition to this, differences in condition or skillsets can reduce the burden of helping further, increasing the times when helping is a worthwhile investment (Jaeggi et al. 2016; Sugiyama and Sugiyama 2003).

When it comes to depression, self-harm, and suicide, the most likely benefits are expected to be long-term changes in how one is treated (Hagen 2003; Hagen et al. 2008; Syme et al. 2016). This can take the form of reconciliations in failing relationships, increased fairness in cooperative relationships, or an increase of resources being directed towards an individual. Of all the benefits, a reduction in physical abuse may be one of the most important, with physical abuse likely being a common occurrence over our evolutionary history based on the high prevalence of it in traditional societies (Stieglitz 2011). 

The benefits of the less costly signals are likely to be more varied due to their low costs allowing smaller benefits to make them worthwhile. These benefits may therefore be more short-term in nature and come from relationships without large conflicts of interest. In regards to crying, a demonstration of need beyond a verbal request may increase the chances of short-term changes in resource allocation or interventions into social conflicts outside of the signaler-target relationship. When it comes to illness symptoms, greater symptomology likely serves to increase the believability that one is ill, thereby increasing the likelihood of social support while also decreasing the risk of reputation hits that might come from the belief that one is faking illness, both of which might be well worth a minor increase in symptomology which might already exist.

Whether long-term or short-term, social support is often classified as emotional or instrumental (material support), with there currently being little empirical evidence on their relative importance (Jacobson 1986). However, if emotions are viewed as motivators to guide behavior towards adaptive outcomes (Nesse 1990; Tooby and Cosmides 2008), it is unlikely emotional support is the end goal. Instead, what appears like emotional support may be the result of achieved support or the promise of future instrumental support. This is especially likely to be the case if such an interaction results in no more need of a signal and, therefore, no more benefit to continuing the negative emotions that had a part in shaping the signal. That being said changes in emotional support may still be informative as they may indicate that an individual believes there is less of a need to signal, either due to support being achieved or indications that signaling need is likely to be ineffective at levels that would result in net benefits.

## Responses to the proposed costly signals of need	

As previously mentioned, the foundational prediction of these signaling theories is that the signaler should often receive benefits after the signal, either in the form of support or renegotiated social relationships. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning (Evraire & Dozois, 2011; Gotlib & Lee, 1989). One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others (Joiner & Metalsky, 1995; Starr & Davila, 2008). Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place (Messer & Gross, 1995; Schwartz et al., 2012). 

Beneficial responses to the depressed are also reported. This is seen with findings that depression reduces the likelihood of aggression within families, particularly in families already experiencing high levels of conflict (Sheeber et al., 2001). There is also limited evidence that depression can result in increased levels of support from others. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together (Hokanson et al., 1986). Likewise, a vignette study found that individuals were more likely to provide advice and support to depressed individuals than anxious individuals despite similar levels of distress (Stephens et al., 1987).

Responses to crying have received more attention, often from a signaling perspective. These responses are generally positive when it comes to infant crying and instrumental outcomes like the amount of care directed towards the infant or its proximity to its parents (Lin & McFatter, 2012). However, crying is also associated with increased distress and other negative emotions in parents and may increase the child’s risk of maltreatment if it is in poor condition (Lin & McFatter, 2012; Soltis, 2004). Responses to adult crying tend to be similar with negative responses typically involving emotions and positive ones involving the likelihood of providing support. For example, individuals faced with crying scenarios in a vignette study experienced more negative emotions relative to controls while also being more likely to provide emotional help and hide negative feelings (Hendriks et al., 2008). Likewise, another experimental study found that while crying increased social support, it came at the cost of being perceived as helpless (Vingerhoets et al., 2016). It is important to note, however, that this trend is not always the case as crying can elicit positive emotions like perceived friendliness and result in negative outward responses like confrontations or mocking at the extreme end (Vingerhoets et al., 2016; Wagner et al., 1997). 

Unfortunately, there are few studies examining how individuals respond to suicide attempt survivors. In those which exist, stigmatization is commonly reported within and outside families (Frey et al. 2017, Scocco et al. 2012). Feelings that survivors are weak, selfish, mentally ill, and antisocial are all commonly reported from a wide array of relationship types (Batterham et al. 2013; Tzeng and Lipson 2004; Frey et al. 2016b).  However, the prevalences of these responses have been found to differ based on relationship type. For example, a study examining the prevalence of stigma found that individuals felt the most stigma from members of their preexisting social networks than mental health providers (Frey et al. 2016a). In these existing networks, participants reported the most stigma among family (57.1% of the time) compared to romantic partners (41.2%) and close friends (28.2%). 

A factor analysis of responses to suicide survivors found that stigma related to suicide attempts could largely be described by three factors: weak, crazy, and distressed (Corrigan et al. 2017). When it existed, prejudice was described by two factors (fear/distrust and anger) and discrimination by three (avoidance, disdain, and coercion). A later vignette study examined these factors across mental health problems and found that participants had higher levels of stigmatization towards imagined characters who attempted suicide (both survived and failed) than depressed or mentally ill characters (Sheehan et al. 2017). In addition to this, the suicide vignettes also resulted in more anger than the other conditions; however, discriminatory behavior was about equal for across the four vignette types.

Although negative responses to suicide survivors are more often studied and reported, positive responses have also been found. Increased social support and changes to important relationships are often reported to follow suicide attempts with some indication that these effects may hold long term. (Stengel 1956). For example, a study of 100 women who survived suicide attempts found that in 75 cases, individuals gained identifiable benefits through the attempt, with 41 individuals benefiting from reconciliations with others (Lukianowicz 1971). As for the type of support received, emotional support seems to be more commonly provided than instrumental support based on initial evidence (Frey et al. 2017); however, it is currently unclear whether one type is more desirable than the other and if this changes with social context.

Unfortunately, most evidence for the benefits of suicide responses in traditional societies is indirect or anecdotal. Unlike Western countries, traditional societies often view suicide attempts as cries for help rather than mental illness, suggesting that help is at least occasionally provided or reliably expected (Shostak 1981;Hidaka 2012). Further evidence comes primarily from the situations in which suicide attempts are made, with suicide attempts often being described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates (Inez HRAF Wilson, Gordon M. (Homicide and suicide among the Joluo of Kenya,  Rafael, The head-hunters of Western Amazonas: the life and culture of the Jibaro Indians of eastern Ecuador and Peru ). Although most reports focus mainly on suicide, there is some indication that the threat of it or a survived attempt can result in increased support, as was the case with the threat of suicide helping a !Kung woman escaping an arranged marriage (Shostak 1981).


## Why responses differ
Due to the connection between depression and social conflict as well as the predicted goal of increased social support, relational level differences are likely to be more important than individual level differences in affecting how others respond to the depressed. This seems to be borne out in the few studies that exist on the topic which show relational level differences being the better predictor of the amount of disclosure suicide survivors were likely to give family members (Fulginiti et al. 2015 in frey et al. 2017), something that is associated with more positive responses from family members and lower depression scores (Frey et al. 2016b). However, it is unclear if it was disclosure itself or being in a more open family that resulted in positive responses.

It is important to note that the relative importance of relational level differences does not mean individual differences are likely to be unimportant. This is especially the case when these differences contribute to differences in one social environment. Although not evidence of responses to need, such a situation is seen with many of the social situations associated with depression being heritable like divorce and family conflict (Kendler and Baker 2007). If positive relationship outcomes are also heritable in a similar way as is reasonable to expect, then individual level differences would also affect how much help is provided in ways that would be hidden by relational variables.

In addition to shaping one’s social environment, individual differences may also result in differences in how effective various signals are, thereby influencing how one is responded to. One area where such a situation might be likely is a sex difference in responses to anger or threats of aggression, as males tend to be stronger than females (Hagen and Rosenström 2016). However, there is currently no evidence to lead to predictions of individual level differences shaping responses to the proposed signals of need.

It is also reasonable to expect that any costly signal of need should be more common with increased likelihoods of help due to the costs involved which which are likely to be compounded by all but the most serious instances of need. Evidence for this may be seen with findings that suggest that some of the proposed costly signals of need are more common in situations where they are more culturally acceptable and where help may be more available. For example, the size of the female biased sex difference in depression was found to differ in relation to gender equity, with more equitable societies experiencing larger gender gaps despite overall lower rates of depression (Hopcroft & Bradley, 2007). The same pattern was seen with crying, with women crying more often in countries with greater gender equality (van Hemert et al. 2011). One possible reason for these relationships is that women in more equal societies experience more crying and depression because they are more likely to receive support, something consistent with the finding that women feel better after crying in countries with relative equality (Becht and Vingerhoets 2002). 

As for explaining how responses differ from signal to signal, the costs of the signals themselves are likely one of the most important variables. This is because, on average, the benefits received need to outweigh the costs of producing the signals, which may be especially high in the costlier proposed signals (Zahavi 1993). Although this seems to suggest that costlier signals should generally result in more favorable responses, it is reasonable to suspect an interaction between signal strength and need whereby stronger signals result in more substantial responses only in times where the signal is proportional to the need. In other words, too strong a signal in times of minor need might send a stronger message of dysfunction than need.

## Hypotheses

Compared to the empirical examinations of the responses of others to the proposed signals of need, the theory is much more advanced and responsible for a number of testable hypotheses. Likely the most foundational is that any costly signal of need should be more common where there is a high degree of private information held by the signaler. This is due to the previously mentioned role of conflicts of interests in selecting for facultative deceptive strategies, which then selected for individuals who are only likely to increase their support of others if their counter-cheating adaptations are satisfied. For this reason, signals should be more common when the target has little information to confirm the signaler is in need and less common when such information exists as the signaling component of any trait or state is designed primarily to provide another organism with information. 

Although instances of private information are likely to result in the highest prevalence of signaling behavior, it is unlikely that they will result in the most helping behavior. Instead, instances of public information of honest need are likely to result in the most help as external information of need is likely a more reliable indicator of need than signals under most circumstances. However, instances of private information are likely to result in more support than public information of cheating with and without signals as even costly signaling attempts may not overcome more reliable information of a lack of need.

Another key hypothesis is that once a costly signal is required to bring about help from others, it is likely that costlier signals will elicit greater benefits all else being equal (Zahavi 1993). This is because of the relationship between the costs of a signal and its honesty which results in costlier signals allowing the target of the signal to be more confident that it is not an attempt at cheating, thereby reducing the probability that the target’s anti-cheating adaptations will prevent him or her from helping. In addition to this, costlier signals, if believed, are also likely to enhance one’s appearance of need further increasing the benefits of helping as long as such help is likely enough to result in future reciprocation, inclusive fitness benefits, or immediate reputation gains. 

Although less examined empirically than the relationship between the cost of a signal and its honesty, it is also reasonable to expect that existing conflict will require costlier signals to be effective. This is mainly due to the likelihood that existing social conflict means larger conflicts of interest are involved, which would therefore require the signaler to be in relatively greater need to be worth helping than in instances with no conflict. High degrees of conflict with the signaler may also increase one’s scrutiny of the signals, which would once again favor costly displays. In addition to fitting signaling theory, such a prediction is also consistent with the previously mentioned predictive relationship between conflict and depression, as depression is an extremely costly state to experience.

## Predictions
In response to the previously mentioned hypotheses, we predict that stronger signals will result in higher levels of both emotional and instrumental social support (i.e. likelihood of providing help or how much money one is willing to give) towards the signaler. Specifically, we predict that the signals will have the following rank order form low to high: no signal (control), verbal request, facial sadness with crying, depression, depression with display of suicidal ideation, and suicide attempt. We also predict that anger will have a positive impact on helping with a similar magnitude to the depression condition. Unlike the hypothesized signals of need, we predict that the schizophrenic break condition will decrease the likelihood of help relative to the control since it indicates cognitive dysfunction.

We also predict that information that the character in the vignette is honest will result in more favorable reactions than the no information condition, which should receive more favorable reactions than conditions where there is information that indicates cheating. A lack of conflict is also expected to result in more favorable reactions than the conflict conditions. 

Although interactions were only briefly examined with depression and the less costly signals here, we also predict several interactions that should hold true for the signals as a whole. We predict that the signals will have the greatest impact on the participant's responses in the private information condition so that the resulting coefficient for signal strength will be greater than in the public information conditions. Similarly, we predict the effect of signal strength to be greater in the conflict condition relative to the no conflict condition. Lastly, we predict that private information will have a greater negative impact on helping in the conflict condition than the no conflict condition.

## Materials and methods

### Vignettes

The vignettes for this study involved participants putting themselves in situations where an imagined sister was requesting a loan of $50,000 from them that was originally saved for their imagined daughter’s college education. Each vignette consisted of two parts, the main storyline which included the conflict and information manipulations and a signal portion that was presented after the participants answered the first set of main outcome questions and the only attention check.

### Factorial Design

The design of this study was a between-subject full factorial design in which three theoretically relevant variables were manipulated: (1) the level of conflict in the relationship, (2) the degree to which one had information about whether or not the person in the vignette is genuinely in need, and (3) the type of emotional signal. 

The amount of conflict within the relationship had a high and a low condition, which was manipulated by describing the participants’ relationship with their sister as “competitive and full of conflict” or “supportive and free of conflict.” The amount of information the participant had on the likelihood the request representing honest need had three conditions: (1) information their sister was misleading them, (2) information their sister was experiencing honest need, and (3) a high degree of private information in which no information about the likelihood of it being an honest request was given. Overall, six types of signals accompanying verbal requests were tested: (1) a verbal request alone, (2) facial sadness with crying, (3) depression, (4) anger, (5) depression with display of suicidal ideation, and (6) a suicide attempt. In addition to these hypothesized signals, two controls were also tested: (1) a no signal control in which the participant attempts to talk to her or his sister but finds that she is not home and (2) a psychotic break condition in which the sister was displaying symptoms consistent with schizophrenia to control for severity alone causing changes in help.

### Sampling

Participants for this study were recruited from Amazon Mechanical Turk and took the survey on Qualtrics. All participants were over 18, located in the United States, had a HIT approval rate of over 95%, and had successfully completed at least 100 HITs as anything less results in approval rates of 100% (Kennedy et al., 2018). To ensure participants were paying attention to the vignette, a 5-question attention check was included which involved questions addressing the topics of manipulations as well as who was making the request and how much money was saved for the daughter. Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed at least 4 of the 5 attention checks were paid $1 for their time.

In total, 1,636 participants passed the required number of attention checks and finished the survey. From this sample of, 68 were excluded for response times under 215 seconds, a time which was selected for being as fast as one could complete the study while paying attention based on trial runs. Of the remaining participants, 119 were also excluded for either having identical GeoIP addresses to at least six other survey completions or an IP address flagged as suspicious of VPS use by the R package rIP. This was due to the reported association between fraudulent, low quality responses to both VPS use and multiple responses (Kennedy et al., 2018). The cutoff of 6 was chosen through the use of regression trees which indicated an attention check failure rate of almost 64% for individuals whose GeoIP matched at least 6 other responses. This left a total sample size of 1,449 for our confirmatory tests.

### Survey

#### Outcome measures

Both emotional and instrumental responses were measured in the survey portion of the study. In a small pilot study, we initially asked the same questions before and after the signal was shown to measure how the signal affected one’s responses. However, due to the prevalence of ceiling and floor effects, the majority of the questions following the signal in this study were revised to ask if and how one’s likelihood of helping or feeling certain emotions changed. This resulted in 22 outcome measures with 11 being asked before the signal and 11 after. To measure these responses, visual analogue scales were used, which allow for finer grained changes than categorical scales and provide a continuum of scores (Klimek et al., 2017). For the pre-signal set of questions, only the ends of each continuum were labeled with some variant of extremely likely and extremely unlikely being used for each measure. For the second set, the center was labelled no change, and the ends of each continuum were labelled as some variant of much more likely and much less likely. Of these 22 variables, 2 are used as a dependent variable in confirmatory analyses here, the self-reported likelihood that participants would help before and after being exposed to the signal.

In addition to this, the amount of money participants were comfortable lending to their imagined sisters was also used as a dependent variable. Unlike the previously described measures, it was repeated identically at both time periods. 

#### Manipulation check

To understand if our manipulations had the desired effect two brief manipulation checks were also included, one before the signal and one following it. The first set of checks involved a question about how close participants felt to the one requesting the help in the imagined scenario and a question about how significant of a request the $50,000 was perceived to be. The second set of checks also had these questions plus one examining what emotional state they thought the imagined signaler was experiencing.

Beyond a way to determine if participants were responding to the study as desired, these questions are also likely meaningful as outcome measures as one’s perception of another’s subjective experience is likely to play a role in their decision making. However, they are only used in exploratory analyses.

#### Demographic Questions

The final part of the survey was a brief demographic questionnaire which examined: (1) the age, (2) sex, (3) state of residence, (4) number of siblings, (5) birth order, (6) number of children, and (7) highest level of education of each participant.

### Analysis

To analyze the effects of conflict and the amount and type of information (hereafter information) on helping behavior, we used a generalized linear model that included conflict and information as independent variables for each initial measure of instrumental support. Although it did not perform too differently from a linear model, a GLM was chosen here and used in all other models due to the prevalence of individuals answering at the extreme ends of the sliders which resulted in error estimates that were not normally distributed.
	
To understand the effects of the signals, we fit a main effects model with no interactions. Here, the manipulated variables included the signals and the amount of help given before the signal, thereby controlling for sampling error while also including the effects of conflict and information that impacted help before the signal. We also fit an interaction model for the two outcome variables in an examination of a smaller subset of signals including depression, crying, and verbal request. This involved the same predictors included in the main effects model in addition to interactions between 1) signal strength and information, 2) conflict and private information, 3) and conflict and signal strength.

## Results

Our total sample size was $N = 1878$. After removal of participants who either did not complete the study or who failed our pre-registered attention checks, our sample size was reduced to $N = `r nrow(signalingdata2018)`$. A further `r sum(signalingdata2018$exclusionopt1)` participants were removed either because they shared an IP address with more than 5 other participants, and were therefore suspected of being a single person, or who finished the survey in less than 215 seconds (the shortest time we were able to complete the survey ourselves) and gave a short, perfunctory response to a free-response item. All such participants were removed prior to any further anayses of the data. Basic demographic information on the remaining $N = `r sum(!signalingdata2018$exclusionopt1)`$ participants (Male = `r sum(d0$Sex == 'Male', na.rm=T)`, Female = `r sum(d0$Sex == 'Female', na.rm=T)`, unspecified = `r sum(is.na(d0$Sex))`) is in Table X. For the number of participants from each US State, see Figure SX.

```{r summarystats}

library(hagenutils)
library(knitr)

demovars <-
  c(
    'Age' = 'Age (years)',
    'years_education' = 'Years of education',
    'Siblings2' = 'Number of siblings',
    'Children' = 'Number of children',
    "Duration (in seconds)" = 'Time to complete survey (seconds)'
  )

kable(
  custom.summarize(d0, demovars),
  caption = 'Summary statistics'
)

```

### Time 1 manipulation checks

At time 1 (T1), participants were randomized into 1 of 2 conflict conditions (conflict vs. support), and 1 of 3 information conditions (cheating, private information, honest), for a total of 6 conditions. Our study was based on the expectation that participants would exhibit less support for the sister in the conflict vs. support conditions, and less support in the cheating vs. private information condition, and more support in the honest vs. private information condition. As we said in our preregistration:

> Amount and type of information:
>
> We predict that information that the character in the vignette is honest will result in more favorable reactions than the no information condition, which should receive more favorable reactions than conditions where there is information that indicates cheating. 
>
> Conflict:
>
> We also predict that the conflict conditions will result in less favorable reactions than the non-conflict condition.
>
> Relationship between emotional and instrumental responses:
>
> In every condition, we predict that emotional responses will track instrumental responses.

To confirm that our manipulations had the desired effects, we analyzed our T1 outcome variables. All 11 T1 outcome variables regarding the sister were strongly positively correlated with each other, except for anger, which was negatively correlated, and sadness about the niece's health problem, which was about the niece and not the sister (and was weakly positively correlated). Among the positively correlated sister variables, correlation coefficients ranged from `r signif(min(cmu), 2)` to `r signif(max(cmu), 2)`, with a mean of `r signif(mean(cmu), 2)`. We therefore conducted a Principle Components Analysis (PCA) of our T1 vars. PC1 accounted for `r signif(100*summary(pcaT1)$importance[2,1], 2)`% of the variance, and all T1 variables loaded equally on this component except for sadness about the sister's daughter's health problem, which loaded primarily on PC2. See Table X and Figure X.

```{r pcaT1}
mat <- pcaT1$rotation[,1:2]
kable(
  mat, 
  digits = 2,
  caption = 'Variables loadings for PCA of  T1 support and emotion variables'
  )
```


```{r biplotT1}
biplotT1
```

We then confirmed that our T1 manipulations had the predicted effect on these outcome variables by fitting a linear regression model of PC1 as a function of the conflict and information conditions. As predicted, PC1 was higher (i.e., more positive toward the sister) in the support vs. conflict condition, and PC1 was also higher in the honest vs. private vs. cheating conditions. See Table X and Figure X.

```{r manipulations, fig.cap="Manipulation checks: the effects of conflict and information conditions on reactions to helping the sister."}
pT1a + pT1b
```

```{r T1Anova}
kable(tidy(Anova(mT1manipulation)), digits = 1, caption="Anova of linear regression of PC1 as a function of the conflict and information conditions.")
```

We predicted the following interaction between conflict and information:

> We also predict that in the private information and cheating vignettes there will be an interaction between the amount of information and levels of conflict such that private information will have a greater negative impact on helping in the conflict condition than the no conflict condition.

Contrary to our prediction, there was no significant interaction between conflict and information in their effects on PC1 at time 1 ($p = `r signif(mT1interaction_anova['conflict:p_info', 'Pr(>F)'], 2)`$).

## Time 2 Signal Effects

### Preregistered tests

Our predicted main effects of signals were:

> We predict that stronger signals  will result in higher levels of both emotional and instrumental social support towards the signaler. Specifically,  we predict that the signals have the following rank order form low to high: 
no signal (control), verbal request, sad face, sad face with crying, depression, depression with display of suicidal ideation, and suicide attempt. We also predict that anger will have a positive impact on helping with a similar magnitude to the depression condition.

This prediction was not upheld. The Control condition (no signal) was ranked relatively high rather than the lowest, Suicide attempt was ranked relatively low rather than the highest, and Anger was consistently ranked lower than Depression. In partial support, however, Depression was consistently higher than Crying, which was consistently higher than Verbal request. See Figure X.

```{r T2signaleffects, fig.cap="Signal effets. The effect of signals on changes in (A) the amount of money the participant is comfortable lending the sister at T2, controlling for the amount at T1 (dotted line indicates mean amount at T1), (B) the likelihood of lending the sister money, (C) participants' perception of the sister's need for money, and (D) the effect of the signal on PC1 of all the outcome variables at T2. Effects are averaged across all time 1 conditions."}
(pT2comfort + pT2lend) / (pT2need + pT2pc1)
```

We also predicted:

> Unlike the hypothesized signals of need, we predict that schizophrenia, which also comes with great fitness costs, will decrease the likelihood of help relative to the control since it indicates cognitive dysfunction.

This prediction was strongly supported: Schizophrenia had far and away the strongest negative impact on participants' emotional attitudes towards, and instrumental support for the sister.

Our predicted interactions were:

> In addition to these main effects, we also predict that the signals will have the greatest impact on the participant's emotional state and likelihood to help when a participant has no information about whether the character is lying or telling the truth. Specifically, we expect the coefficient for signal strength to increase in the private information condition relative to the public information conditions.
>
> We also predict that when the most severe signals, i.e., depression and suicidality,  are paired with information that the character in the vignette is likely to be lying, responses will be more aversive than when the signal is weaker.

```{r pinfoInteractions, fig.cap=""}
(p_comfort_signal_pinfo + p_lend_signal_pinfo + scale_y_continuous(limits = c(-40, 20)))/(p_pc1_signal_pinfo + p_money_signal_pinfo + scale_y_continuous(limits = c(-40, 20)))
```

```{r conflictInteractions. fig.cap=""}
(p_comfort_signal_conflict + p_lend_signal_conflict + scale_y_continuous(limits = c(-40, 20)))/(p_pc1_signal_conflict + p_money_signal_conflict + scale_y_continuous(limits = c(-40, 20)))
```


## Discussion

## Limitations
The key limitation of this study is that evidence for increased helping in response to stronger signals of need, while suggestive of adaptations for helping important others in need, is not evidence that the symptoms of the proposed signals evolved or were coopted for this purpose. Due to the benefits of inclusive fitness or indirect reciprocity, it is likely that some situations may result in help without the need of a signal, especially if there is little private information held by the one in need. However, the existence of helping adaptations is a prerequisite for the evolution of any signal of need as there first must be benefits to deception before there can be selection for adaptations to detect cheaters which would then lead adaptations meant to overcome the anti-cheating adaptations of others.
	However, cheating strategies to extract more aid than another would want to give are likely to go beyond simply being a prerequisite and may be a strong enough selection pressure to ensure that the symptoms relevant to our anti-cheater adaptations are rarely unaltered by a signaling component. This is especially likely when the signal takes the form of an increase in symptoms likely to already exist due to one’s condition. For example, a sick individual might increase the severity of one’s external symptoms beyond what is ideal for immune defense to increase the likelihood of help from others (Steinkopf, 2015; Tiokhin, 2016). This has mainly been explored as an adaptive explanation for placebo responses; however, it also has the potential to be the case with depression as the symptoms are largely identical to the types of symptoms seen in normal, differing mainly in severity and duration (Nesse, 1999; Wakefield & First, 2012).
	
# Supplementary information

```{r numbersbystate, fig.height=10}
dotchart(sort(table(signalingdata2018$State)))
```

